Can we constrain galaxy assembly bias with group and clustering statistics of SDSS galaxies?

### 2016-09-15 text/ms.tex

\title{Can we constrain galaxy assembly bias with group and clustering statistics of SDSS galaxies?}
\author{Mohammadjavad Vakili  \altaffilmark{1,$\dagger$}, ChangHoon~Hahn \altaffilmark{1}, Unspecified contributors \altaffilmark{1}}
\altaffiltext{$\dagger$}{\url{mjvakili@nyu.edu}}
\altaffiltext{1}{Center for Cosmology and Particle Physics, 
                 Department of Physics, New York University, 4 Washington Pl., New York, NY} 
\begin{abstract}

The standard halo occupation modeling approach in large scale structure 
studies assumes that the halo mass alone suffices in characterizing the 
galaxy-halo connection. Galaxy assembly bias, the hypothesis that properties beyond 
halo mass influences the galaxy-halo connection, can potentially affect the 
statistical summaries of galaxy populations. The degree to which, assembly bias can 
can be constrained with the statistical distribution of galaxies remains largely 
untested. In this investigation, we make use of the recently proposed assembly bias decoration 
of the HOD modeling in order to find constraints on the strength of the assembly bias from the most up-to-date and publicly available measurements of the group multiplicity function and the projected two-point correlation function of the Volume-limited luminosity threshold sample of SDSS 
galaxies. We find that the available GMF measurements cannot constrain galaxy assembly bias in any of the Luminosity thresholds. On the other hand, we find non-zero constraints on the satellite galaxy assembly bias for the $M_{r}<-18$ and $M_{r}<-19$ samples, nonzero constraints on the central galaxy assembly bias for the $M_{r}<-20$ samples, and constraints consistent with zero for the assembly bias of galaxies in the $M_{r}<-21$ sample. Furthermore, we provide a model comparison between the standard HOD approach and the decorated HOD in modeling of the GMF and the projected 2PCF measurements.

\end{abstract}

>>>>

\author{Mohammadjavad Vakili\altaffilmark{1}, David~W.~Hogg\altaffilmark{1,2,3,4}, and ChangHoon~Hahn\altaffilmark{1}}
\affil{Center for Cosmology and Particle Physics, New York University, 4 Washington Pl, New York, NY, 10003}
\email{mjvakili@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, New York University, 4 Washington Pl, New York, NY, 10003}
\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, New York, NY}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, New York, NY}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\begin{abstract}
Analyses of the $N$-body simulations have demonstrated that the clustering of dark matter halos depend on the halo properties beyond mass such as halo concentration. Standard halo occupation distribution (HOD) modeling approach in large scale structure studies assumes that the halo mass alone is sufficient in characterizing the galaxy-halo connection. 
Galaxy assembly bias hypothesizes that properties beyond halo mass influence the galaxy-halo connection, and can lead to systematic effects on the modeling of galaxy clustering.

Using the Small MultiDark Planck high resolution $N$- body simulation and the measurements of the projected two-point correlation function as well as the number density of the DR7 main galaxy sample of the Sloan Digital Sky Survey, we constrain the concentration-dependence of HOD. This concentration dependence is modeled within the framework of the proposed decorated HOD model.

For fainter samples, we find non-zero albeit poor constraints on the strength of the assembly bias. Furthermore, we find that the assembly bias remains largely unconstrained in the brighter galaxy samples that live in massive dark matter halos.

By comparison between the 2PCF measured from the mock galaxy catalogs using our best-fit model and the randomized galaxy catalogs in which the assembly bias is erased, we note that the effect of assembly bias on galaxy clustering is qualitatively similar to its effect on galaxy catalogs created from a certain class of subhalo abundance matching models. 
\end{abstract}
\keywords{Cosmology: large-scale-structure, galaxy: galaxies}
\end{abstract}
